Synthesizing Quotient Lenses

### 2018-03-04 papers/QuotientLenses/main.tex
edit
minor fixes

% % Copyright information % Supplied to authors (based on authors' rights
% management selection; % see authors.acm.org) by publisher for camera-ready
% submission; % use 'none' for review submission.
\setcopyright{none}
\usepackage{amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz,
array, mathtools, bussproofs, stmaryrd, enumitem, stackengine, subcaption}
\captionsetup{compatibility=false}
\usepackage{relsize}
\usepackage{listings}
\usepackage{proof}

@@

\newcommand{\Name}{Optometrist}

@@

\newcommand{\OpticianRuntime}{\textbf{OO}}
\newcommand{\SystemOnOptician}{\textbf{SO}}
\newcommand{\SystemOnBenchmarks}{\textbf{SS}}
\newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}

@@

We have integrated QREs and quotient lens synthesis into Boomerang allowing QREs
and synthesis tasks to be expressed alongside existing Boomerang expressions.

@@

Our evaluations serve to answer the following questions:
\begin{enumerate}
  \item Does synthesizing quotient lenses from QREs permit for an easier
  development process than existing approaches?

  \item Are QREs sufficiently expressive that more lenses can be synthesized
  without modifications to the format?

  \item Is our synthesis algorithm competitive with existing lens synthesis
  tools?
\end{enumerate}

@@

\begin{figure}
\includegraphics{generated-graphs/asts.eps}
\caption{Count of benchmark programs definable using a given AST count. We find
that it takes far fewer AST nodes to define benchmark lenses in \Name{} than
with Optician or without synthesis. }
\label{fig:asts}
\end{figure}

To test if our tool eases the development process, we compare the following
number of AST nodes.
%
\begin{itemize}
  \item[\QRESize{}] \QRESize{} is the number of AST nodes in the QRE
  specifications involved in QRE lens synthesis task.
  \item[\CanonizerAndSpecSize{}] \CanonizerAndSpecSize{} is the number of
  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
  for each QRE used in QRE lens synthesis task.
  \item[\LensAndSpecSize{}] \LensAndSpecSize{} is the number of AST nodes
  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
\end{itemize}

@@

The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
benchmarks without QREs.  This is only an approximation, as both $W(q)$ and
$\canonizer(q)$ are automatically generated from the QRE.  In our
experience, these automatically generated regular expressions and canonizers are
the ones we would naturally write.

The value of \LensAndSpecSize{} represents the difficulty in writing our benchmarks
with neither QREs nor synthesis.  This is also an approximation, as $W(q)$ and
$\canonizer(q)$ and the bijective lens are all automatically generated.
We have found that the generated bijective lens is fairly similar in size to the
one we would manually write, though it is typically less modular.

@@

\subsection{Increase of Expressivity}

Optician was only able to synthesize fully bijective data transformations.
\Name{} loosens that restriction, and is able to synthesize data which is
bijective when put in canonical form.
To determine the increase of expressivity, we analyzed:
\begin{enumerate}
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective.
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective when put in canonical form.
\end{enumerate}

We found that 10 of Opticians 39 benchmarks required modifications to be
bijective. We found that 4 of Optician's 39
benchmarks required modifications to be in bijective correspondence when in
canonical form.  Furthermore, we found that all of the data.gov examples
required no modifications to be in canonical form.  While we did not put the
data.gov examples into a form suitable for Optician, we found only 4 of
the 8 data.gov examples did not require canonization, so making suitable
for Optician would have required a modification of 4 of the data.gov
formats.

\subsection{Competitive Performance}

\begin{figure}
\centering
\begin{subfigure}[b]{.49\textwidth}
\centering
\includegraphics{generated-graphs/times_opt}
\caption{}
\label{subfig:lenssize}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}
\includegraphics{generated-graphs/times_new.eps}
\caption{}
\label{subfig:examplesused}
\end{subfigure}
\caption{Runtimes of Optician and \Name{}.
In (a), we run Optician and \Name{} on Optician's benchmarks.  We find that
there is only a negligible difference incurred by Optometrist's overhead.
In (b), we run Optometrist on its benchmark suite.  We find that it is able
to synthesize all quotient lenses in under 20 seconds, and typically
finished in under 5 seconds. }
\label{fig:times}
\end{figure}

While we expanded on the Optician system, we found it important to validate
performance did not degrade in this expansion.  To validate this, we tested:
\begin{itemize}
  \item[\OpticianRuntime{}] Optician runs on its own benchmarks.
  \item[\SystemOnOptician{}] \Name{} runs on Optician's benchmarks.
  \item[\SystemOnBenchmarks{}] \Name{} runs on its own benchmarks.
\end{itemize}

We summarize the results of these experiments in Figure~\ref{fig:times}.
\Name{} was
able to synthesize all of Optician's benchmarks at a speed competitive with
Optician.  There was additional overhead in calculating $W(q)$ and $K(q)$,
resulting in a very slight decrease in performance.

Also notable was one of the data.gov examples, which converted demographic
statistics by zip code from xml form into rdf form.  This merely took longer because it was
an exceptionally large format, converting over 40 fields.  After adding in additional
fields to make this format bijective, it still took over 15 seconds to complete.

>>>>

% % Copyright information % Supplied to authors (based on authors' rights
% management selection; % see authors.acm.org) by publisher for camera-ready
% submission; % use 'none' for review submission.
\setcopyright{none}
\usepackage{amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz,
array, mathtools, bussproofs, stmaryrd, enumitem, stackengine, subcaption}
\captionsetup{compatibility=false}
\usepackage{relsize}
\usepackage{listings}
\usepackage{proof}
\usepackage{xspace}

@@

\newcommand{\Name}{Optometrist\xspace}

@@

\newcommand{\OpticianRuntime}{\textbf{OO}}
\newcommand{\SystemOnOptician}{\textbf{QO}}
\newcommand{\SystemOnBenchmarks}{\textbf{QQ}}
\newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}

@@

We have integrated QREs and quotient lens synthesis into Boomerang. With this
integration, QREs and quotient lens synthesis tasks can be expressed alongside
existing Boomerang terms.

@@

Our evaluations serve to answer the following questions:
\begin{enumerate}
\item Does synthesizing quotient lenses from QREs permit an easier development
  process than existing approaches?

\item Does the increase in expressivity gained from synthesizing from QREs
  permit synthesis of more useful lenses than existing approaches.

\item Does our tool synthesize lenses fast enough to be used as part of a
  standard development process.
\end{enumerate}

@@

\begin{figure}[t]
\includegraphics{generated-graphs/asts.eps}
\caption{Count of benchmark programs definable using a given AST count. We find
  that it takes far fewer AST nodes to define benchmark lenses using QRE
  synthesis than with Optician without QRE synthesis or without synthesis. }
\label{fig:asts}
\end{figure}

To test if our tool eases the development process, we compare the following
number of AST nodes.
%
\begin{itemize}
  \item[\QRESize{}] is the number of AST nodes in the QRE
  specifications involved in QRE lens synthesis task.
  \item[\CanonizerAndSpecSize{}] is the number of
  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
  for each QRE used in QRE lens synthesis task.
  \item[\LensAndSpecSize{}] is the number of AST nodes
  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
\end{itemize}

@@

The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
benchmarks within the Optician system. This is only an approximation, as both
$W(q)$ and $\canonizer(q)$ are automatically generated from the QRE, instead of
written by a human.

The value of \LensAndSpecSize{} represents the difficulty in writing our
benchmarks in Boomerang with neither QREs nor synthesis. This is also an
approximation, as $W(q)$ and $\canonizer(q)$ and the bijective lens are all
automatically generated.

@@

\subsection{Increasing Expressivity}

Optician was only able to synthesize fully bijective data transformations.
Synthesizing from QREs loosens that restriction, making it possible to
synthesize data which is bijective when put in canonical form. To determine the
increase of expressivity, we analyzed:
\begin{enumerate}
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective.
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective when put in canonical form.
\end{enumerate}

We found that 10 of Opticians 39 benchmarks required modifications to be
bijective. We found that 4 of Optician's 39 benchmarks required modifications to
be in bijective correspondence when in canonical form. Furthermore, we found
that all of the data.gov examples required no modifications to be in bijective
correspondence when in canonical form. While we did not put the data.gov
examples into a form suitable for Optician, we found only 4 of the 8 data.gov
examples did not require canonization, so making suitable for Optician would
have required a modification of 4 of the data.gov formats.

\subsection{Maintaining Competitive Performance}

\begin{figure}[t]
\centering
\begin{subfigure}[b]{.49\textwidth}
\centering
\includegraphics{generated-graphs/times_opt}
\caption{}
\label{subfig:lenssize}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}
\includegraphics{generated-graphs/times_new.eps}
\caption{}
\label{subfig:examplesused}
\end{subfigure}
\caption{Runtimes of Optician and our system.
In (a), we run Optician and our system on Optician's benchmarks.  We find that
there is only a negligible difference incurred by Optometrist's overhead.
In (b), we run Optometrist on its benchmark suite.  We find that it is able
to synthesize all quotient lenses in under 20 seconds, and typically
finished in under 5 seconds. }
\label{fig:times}
\end{figure}

To be used in a development environment, we found it important to validate
performance did not significantly degrade from Optician. To validate this, we
tested the runtimes when:
\begin{itemize}
  \item[\OpticianRuntime{}] Optician runs on its own benchmarks.
  \item[\SystemOnOptician{}] Our system runs on Optician's benchmarks.
  \item[\SystemOnBenchmarks{}] Our system runs on its own benchmarks.
\end{itemize}

We summarize the results of these experiments in Figure~\ref{fig:times}.
Our system was
able to synthesize all of Optician's benchmarks at a speed competitive with
Optician.  There was additional overhead in calculating $W(q)$ and $K(q)$,
resulting in a very slight decrease in performance.

Also notable was one of the data.gov examples, which converted demographic
statistics by zip code from xml form into rdf form. This merely took longer
because it was an exceptionally large format, converting over 40 fields. We
added in additional fields to make this format suitable for Optician, which
took over 15 seconds synthesize the modified lens.
